Rawls For Electricity Planning Models

### 2024-12-02 Paper_Endogenous.tex

Previous work has included co-benefits involved with reducing electricity generation from fossil fuels in the objective function \citet{Sergi2020OptimizingBenefits}.
or projected how the electricity transmission might affect the distributional consequences of air pollution exposure \citep{Goforth2022AirStrategies}. However, in this paper, rather than modifying the objective function to include health effects, or use model outputs to project health changes, we endogenize pollution exposure within the electricity model in a way that explicitly considers distributional effects.  This allows us to use the model to produce results that don't unfairly burden certain groups with disproportionate pollution exposure.  It also can be used to answer questions such as "How much can pollution exposure be equitably reduced for $\$100$ billion?"

In this paper, we use

\subsection{Power Sector Modelling Inputs}
Stuff about PowerGenome.

\section{Methods}

@@

In this paper, we outline a method to use least-cost electricity planning models to design power systems that produce equitable distributions of air pollution exposure.  We pull inspiration from Rawls and formulate our model as a min-max problem where the social planner minimizes the maximum pollution exposure subject to a cost constraint, rather than minimizing costs.  Running multiple iterations of the model at different cost constraints allows us to understand, how much can we equitably improve air quality and at what costs.
\begin{figure}
    \begin{center}
        \begin{tikzpicture}[node distance=2cm]
        \node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
        • Constraints}};

@@

\subsection{Air Pollution Modelling}
The end product necessary to endogenize air pollution within electricity planning models are a set of coefficients that describe the marginal effect of a unit of electricity generation on population-weighted mean pollution exposure for different racial groups (though you could also apply the same logic and methodology to differing socio-economic groups more broadly).

To do this we employ a reduced-form source receptor matrix that that describe the marginal effect of a unit of pollution on downwind concentration of pollution exposure called the InMap Source Receptor Matrix (ISRM).  InMap is a reduced-form emissions to exposure model that uses annual-average parameters to reduce computation complexity relative to a time-resolved chemical transport model. InMap is designed to model annual average PM2.5 pollution exposure, and takes as inputs annual emissions of volatile organic compounds (VOCs), nitrous oxides (NOx), ammonia (NH3), sulfur oxides (SOx), and primary PM2.5.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{}
    \caption{Caption}
    \label{fig:enter-label}
\end{figure}

@@

\section{Conclusion}

>>>>

Previous work has included co-benefits involved with reducing electricity generation from fossil fuels in the objective function \citet{Sergi2020OptimizingBenefits}.
or projected how the electricity transmission might affect the distributional consequences of air pollution exposure \citep{Goforth2022AirStrategies}.

@@

However, in this paper, rather than modifying the objective function to include health effects, or use model outputs to project health changes, we endogenize pollution exposure within the electricity model in a way that explicitly considers distributional effects.  We pull inspiration from Rawls and formulate our model as a min-max problem where the social planner minimizes the maximum pollution exposure subject to a cost constraint, rather than minimizing costs.  Running multiple iterations of the model at different cost constraints allows us to understand how much can we equitably improve air quality and at what costs.

This allows us to use the model to produce results that don't unfairly burden certain groups with disproportionate pollution exposure.  It also can be used to answer questions such as "How much can pollution exposure be equitably reduced for $\$100$ billion?"

This paper makes an important contribution to the environmental justice literature by ...

\subsection{Power Sector Modelling Inputs}
Stuff about PowerGenome?

\section{Methods}
\subsection{Overview}

\begin{figure}
    \begin{center}
        \begin{tikzpicture}[node distance=2cm]
        \node (start) [start, text width=5cm] {\shortstack[l]{Inputs\\• Electricity Demand \\ • Costs of New Generation \\ • Generation Characteristics\\
        • Constraints}};

@@

\subsection{Air Pollution Modelling}
The end product necessary to endogenize air pollution within electricity planning models are a set of coefficients that describe the marginal effect of a unit of electricity generation on population-weighted mean pollution exposure for different racial groups (though you could also apply the same logic and methodology to differing socio-economic groups more broadly, or more fine slices of the population to more closely approximate a true min-max).

Eulerian chemical transportation models are computationally expensive and time consuming, so it's infeasible to embed this type of model within SWITCH.  Rather we use the InMAP Source-Receptor Matrix (ISRM) which is a reduced-form source-receptor matrix that describes the marginal effect of a unit of pollution on downwind concentrations of pollution exposure.

InMap itself is a reduced-form emissions-to-exposure model that uses annual-average parameters to reduce computation complexity relative to a time-resolved chemical transport model. InMap is designed to model annual average PM2.5 pollution exposure, and takes as inputs the annual emissions of volatile organic compounds (VOCs), nitrous oxides (NOx), ammonia (NH3), sulfur oxides (SOx), and primary PM2.5 \citep{Tessum2017InMAP:Interventions}.

The ISRM was developed to simplify InMAP even further \citep{Goodkind2019Fine-scaleEmissions}.  Initially, the ISRM was developed to describe the marginal damages from individual sources.  We simplify the ISRM even further by collapsing it so that, for each source, we save the marginal effect of one MWh of generation on \textit{population-weighted average exposure} by race.

\subsection{Switch}

@@

\begin{figure}
    \centering
    \subcaptionbox{2027}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/Dispatch_Relative_by_scenario_2027.png}}
    \subcaptionbox{2030}{\includegraphics[width=0.45\textwidth]{Figures/EndogenousResults/base_short/Dispatch_Relative_by_scenario_2030.png}}
    \caption{Generation by Technology by Scenario}
\end{figure}

@@

\section{Conclusion}
In this paper, we demonstrate how to endogenize air pollution exposure into a least-cost electricity planning model to create equitable reductions in air pollution burdens.

While this method provides a map of how to produce equitable reductions in air pollution, it does not solve for what policies will get
